---
title: "Homework 1 - Bumpus sparrows"
author: "Batul Khambata"
date: "02/04/2023"
output: html_document
---
### Importing data set
```{r}
library(readr)
Bumpus_sparrows = read_csv("/Users/batul/Downloads/Bumpus_sparrows.csv")
Bumpus_sparrows
sparr = Bumpus_sparrows
sparr

#Inference
#Correlation
#Correlation between Keel Sernum and alar extent is higher in bumpus sparrows that didn't survive comapred to sparrows that survived

```
### Checking if there are null values in the dataset
```{r}
any(is.na(sparr))
```
### Splitting data depending on the Survivors and Non Survivors
```{r}
split_list = split(sparr, f = factor(sparr$Survivorship))
Not_Survived = split_list[[1]]
Survived = split_list[[2]]
```
### Dropping the first column
```{r}
NSur = Not_Survived[,-1]
Sur=Survived[,-1]
```
### Calculating the Mean of each column for Survivors and Remnants
```{r}
colMeans(NSur)
colMeans(Sur)
```
### Calculating the Covariance of each column for Survivors and Remnants
```{r}
NS_cov = cov(NSur, y=NULL, method = "pearson")
NS_cov

S_cov = cov(Sur, y=NULL, method = "pearson")
S_cov
```
* Inference
### Calculating the Correlation of each column for Survivors and Remnants
```{r}
NS_corr = cor(NSur, y=NULL,  method = "pearson")
NS_corr
S_corr = cor(Sur, y=NULL,  method = "pearson")
S_corr
```
### Plotting a matrix for Correlation for Survivors and Remnants
```{r}
library(ggcorrplot)
ggcorrplot(NS_corr, lab = TRUE)
ggcorrplot(S_corr, lab = TRUE)
```
```{r}


```

### Covariance Matrix
```{r}
cov_matrix <- cov(Not_Survived$L_keel_sternum, Not_Survived$Total_length)
plot(Not_Survived$L_keel_sternum, Not_Survived$Total_length, xlab="Variable 1", ylab="Variable 2", main="Not Survived Sparrows Covariance Plot")

cov_matrix <- cov(Survived$L_keel_sternum, Survived$Total_length)
plot(Survived$L_keel_sternum, Survived$Total_length, xlab="Variable 1", ylab="Variable 2", main="Survived Sparrows Covariance Plot")
```
### Box Plot
```{r}
ggplot(sparr,aes(y=Alar_extent, fill=Survivorship))+geom_boxplot()+ labs(title='Comparison Bewtween Alax extent in Bumpus Sparrows')

ggplot(sparr,aes(y=L_beak_head, fill=Survivorship))+geom_boxplot()+ labs(title='Comparison between length of beak and head in Bumpus Sparrows')

ggplot(sparr,aes(y=L_humerous, fill=Survivorship))+geom_boxplot()+ labs(title='Comparison between length of humerous in Bumpus Sparrows')

ggplot(sparr,aes(y=L_keel_sternum, fill=Survivorship))+geom_boxplot()+ labs(title='Comparison between length of keel and sternum in Bumpus Sparrows')

ggplot(sparr,aes(y=Total_length, fill=Survivorship))+geom_boxplot()+ labs(title='Comparison between total length of Bumpus Sparrows')
```

### Inference

* This boxplot shows the difference between total length in sparrows that survived versus sparrows that died. Overall, sparrows that survived were shorted than dead sparrows.
* We can also see from the correlation matrix that there is a stronger relationship between alar extent and keel sternum for sparrows that dies and very weak relationship between alar extent and keel sternum for sparrows that survived.
* We can conclude that higher sparrow body length results in less chances of survival of the Bumpus Sparrows. Similarly sparrows that are shorter have higher chances of survival.




